### 2018-06-03 Sections/optimising_the_world.tex
Search tree sec and fig

Every aspect of our modern world is heavily optimised to improve efficiency, affordability, profitability, or any other kind of desirability. Even long before the digital revolution, which enabled the widespread implementation of modern optimisation-theoretic techniques, humans have always striven to make life as easy as possible -- the least effort for the greatest possible reward. Our modern application of optimisation theory is merely a logical extension of this to the many other facets of the 21st century economy that now exist.

@@

It is well known that the most complex such satisfiability problems are \textbf{NP}-complete\index{\textbf{NP}-complete} in general, with no known efficient classical solutions (in fact even very trivial constructions of satisfiability problems are already \textbf{NP}-complete, e.g Fig.~\ref{fig:3SAT}).

@@

We are once again in a situation where we are overwhelmed with exponentially growing combinatorics to try and minimise the countless possible schedule conflicts that can occur. It is obvious upon inspection that finding a global optimum to this problem will not be possible via independent local optimisations.

@@

@@

The modern economic infrastructure for the production and distribution of goods is built upon supply chains -- complex, interdependent networks of the exchange of resources between entities as goods pass through their many stages of production before reaching the consumer. Each exchange of resources will typically be subject to its own constraints, such as strict time of arrival demands. Effectively this yields a massive instance of a complex satisfiability problem, where all, or as many as possible of the demands of the units in the chain must be simultaneously met.

@@

Classical techniques for attacking optimisation problems such as the ones presented above typically come in two flavours:
\begin{itemize}
	\item Heuristics: efficient algorithms that find suboptimal, but satisfactory \textit{approximations} to the solution to a problem. These algorithms are fast and can be efficiently implemented classically, but can have highly variable accuracy in the solution they provide.
	\item Brute-force: use raw computational power to exhaustively work through all the combinatorics of a problem. This is guaranteed to find the optimal solution, but is typically extremely slow and limited to small instances of the problem.
\end{itemize}

@@

However, such problems can be quadratically enhanced in runtime by cunningly employing Grover's algorithm. To see this, note that all \textbf{NP}-complete problems can be efficiently mapped to one another with polynomial resource overhead (see Fig.~\ref{fig:complexity_classes}). Thus, we can restrict ourselves to considering the satisfiability problems\index{Satisfiability problems} discussed above -- the archetypal \textbf{NP}-complete problems. An example of the 3-\textsc{SAT} problem\index{3-SAT problem}, which is \textbf{NP}-complete, is shown in Fig.~\ref{fig:3SAT}.

\begin{figure}[!htb]
\includegraphics[width=0.3\textwidth]{3SAT}
\caption{Digital circuit for an instance of the 3-\textsc{SAT} problem, with 4 clauses acting on input variables $\{x_i\}$. Each of the OR gates is input with some combination of 3 of the input bits or their compliments (i.e with a NOT gate). Each of these is referred to as a `clause', of which there are 4 in this example, but could be any number in general. The final AND gate requires that all clauses be simultaneously satisfied in order to yield a final output of `1'. The goal of the problem is to find an input bit-string $x$ that yields an output of `1'. In general, this may require exhaustively searching over the entire space of input states via brute-force, which exhibits time-complexity exponential in the length of the bit-string. This problem is proven to be \textbf{NP}-complete. Note that the similarly-defined 2-\textsc{SAT}\index{2-SAT problem} problem (i.e clauses each contain 2 input bits) is \textbf{P}, making 3-\textsc{SAT} the simplest model to consider in the study of \textbf{NP}-complete problems. For this reason 3-\textsc{SAT} is often used as the computational model when studying \textbf{NP}-complete problems.} \label{fig:3SAT}	
\end{figure}

By defining an oracle\index{Oracle} that implements a polynomial-time algorithm on $n$ qubits, a Grover search over the input space of $2^n$ configurations will determine the satisfying input to the oracle for a given desired output, which acts as the tagged element within the search algorithm. The Grover search yields a quadratic speedup for this search compared to a brute-force classical search, therefore requiring only $O(2^{n/2})$ oracle calls. While this is short of the exponential speedup one might hope for, a quadratic speedup can nonetheless be very significant for large problem instances, where even minor improvements could be of enormous value.

@@

Some problems in even harder classes than \textbf{NP}-complete can in some instances be \textit{approximated} using the same approach. The key to solving such problems is to define an oracle\index{Oracle} that attributes a \textit{score}\index{Score} to a given input, rather than a yes/no answer to satisfiability, and answers `yes' or `no' depending on whether that score is above some defined threshold. As an illustrative example, consider the optimisation of, say, a complex traffic network, where the goal is to maximise flow through the network. Then we might define our score to be some flow metric for the network's graph.

We then apply a Grover search repeatedly, each time incrementing this threshold until the algorithm outputs `no'. Then we know that the last input had the highest score. The reason this approach is \textit{approximate} rather than \textit{exact} is that defining such a score-oracle mightn't be always efficiently implemented, or maybe it mightn't make sense at all to define score measures for a given problem.

An intuitive example of a problem taken from a far harder complexity class than \textbf{NP}-complete, that can be approximately optimised using the Grover approach is the game of Chess\index{Chess}. Generalisation of the rules of Chess to an \mbox{$n\times n$} board yields a game proven to reside in \textbf{EXP}-complete\index{\textbf{EXP}-complete} -- an \textit{extremely} hard class.

How do current classical computers play chess? They construct a search tree, which is searched exhaustively up to a given depth (bounded by computational resources), and assign a score to each board position, indicating the relative strength of its position. The branch in the tree with the highest scores determines the next move.

@@

\subsection{\textbf{BQP} optimisation problems}\index{\textbf{BQP} optimisation problems}

The optimisation problems discussed until now have been satisfiability problems residing in \textbf{NP}-complete, limited to only quadratic quantum enhancement. Are there any other optimisation problems that natively reside in \textbf{BQP}, and therefore inherently offer exponential quantum speedup? There are, although the examples are far less ubiquitous than the plethora of satisfiability-based optimisation problems.

\comment{To do!}

@@

While the dream-goals of perfectly optimised resource allocation and perfect economic efficiency are fantasy, we should not underestimate the impact that even modest enhancements in the efficiency of large-scale economic systems could have. While quantum computers will never perfectly solve humanity's immense resource allocation problems, even the modest improvements they offer could be of immense value to the world economy, with enormous implications monetarily, socially, and environmentally. Indeed, optimisation problems are so ubiquitous that pursuing quantum-enhanced optimisation will be of central importance in the quantum era.

>>>>

Every aspect of our modern world is heavily optimised to improve efficiency, affordability, profitability, and every other kind of desirability. Even long before the digital revolution, which enabled the widespread implementation of modern optimisation-theoretic techniques, humans have always striven to make life as easy as possible -- the least effort for the greatest possible reward. Our modern application of optimisation theory is merely a logical extension of this to the many other facets of the 21st century economy that now exist.

@@

It is well known that the most complex such satisfiability problems are \textbf{NP}-complete\index{\textbf{NP}-complete} in general, with no known efficient classical solutions (in fact even very trivial constructions of satisfiability problems are already \textbf{NP}-complete, e.g see Fig.~\ref{fig:3SAT}).

@@

We are once again in a situation where we are overwhelmed with exponentially growing combinatorics to try and minimise the countless possible schedule conflicts that can occur. It is obvious upon inspection that finding a global optimum to this problem will not be possible via independent local optimisations, which do not accommodate for competing interdependencies.

@@

Of course, these ideas don't just apply to government budgets, but also budgets at smaller scales, say within a large corporation with many distinct spending programs.

@@

The modern economic infrastructure for the production and distribution of goods is built upon supply chains -- complex, interdependent networks of the exchange of resources between entities as goods pass through their many stages of production before reaching the consumer. Each exchange of resources will typically be subject to its own constraints, such as strict time of arrival demands. Effectively this yields a massive instance of a complex satisfiability problem, where all, or as many as possible of the demands of the units in the chain must be simultaneously met. Conceivably, a single unsatisfied demand may break the functionality of the entire supply chain!

@@

Classical techniques for attacking optimisation problems, such as the ones presented above, typically come in two flavours:
\begin{itemize}
	\item Heuristics: efficient algorithms that find suboptimal, but satisfactory \textit{approximations} to the solution to a problem. These algorithms are fast and can be efficiently implemented classically, but can have highly variable accuracy in the solution they provide.
	\item Brute-force: use raw computational power to exhaustively work through all the combinatorics of a problem. This is guaranteed to find the optimal solution, but is typically extremely slow and limited to small instances of the problem.
\end{itemize}

@@

However, such problems can be \textit{quadratically} enhanced in runtime by cunningly employing Grover's search algorithm (Sec.~\ref{sec:quantum_search}). To see this, note that all \textbf{NP}-complete problems can be efficiently mapped to one another with polynomial resource overhead (see Fig.~\ref{fig:complexity_classes}). Thus, we can restrict ourselves to considering the satisfiability problems\index{Satisfiability problems} discussed above -- the archetypal \textbf{NP}-complete problems. An example of the 3-\textsc{SAT} problem\index{3-SAT problem}, which is \textbf{NP}-complete, is shown in Fig.~\ref{fig:3SAT}.

\begin{figure}[!htb]
\includegraphics[width=0.3\textwidth]{3SAT}
\caption{Digital circuit for an instance of the 3-\textsc{SAT} problem, with 4 clauses acting on input variables $\{x_i\}$. Each of the OR gates is input with some combination of 3 of the input bits or their compliments (i.e with a NOT gate). Each of these is referred to as a `clause', of which there are 4 in this example, but could be any number in general. The final AND gate requires that all clauses be simultaneously satisfied in order to yield a final output of `1'. The goal of the problem is to find an input bit-string $x$ that yields an output of `1'. In general, this may require exhaustively searching over the entire space of input states via brute-force, which exhibits time-complexity exponential in the length of the bit-string and associated number of clauses. This problem is proven to be \textbf{NP}-complete. Note that the similarly-defined 2-\textsc{SAT}\index{2-SAT problem} problem (i.e clauses each contain 2 input bits) is \textbf{P}, making 3-\textsc{SAT} the simplest model to consider in the study of \textbf{NP}-complete problems. For this reason 3-\textsc{SAT} is often used as the computational model when studying \textbf{NP}-complete problems.} \label{fig:3SAT}	
\end{figure}

By defining an oracle\index{Oracle} that implements a polynomial-time algorithm on $n$ qubits, a Grover search over the input space of $O(2^n)$ configurations will determine the satisfying input to the oracle for a given desired output, which acts as the tagged element within the search algorithm. The Grover search yields a quadratic speedup for this search compared to a brute-force classical search, therefore requiring only $O(2^{n/2})$ oracle calls. While this is short of the exponential speedup one might hope for, a quadratic speedup can nonetheless be very significant for large problem instances, where even minor improvements could be of enormous value.

@@

Some problems in even harder classes than \textbf{NP}-complete can in some instances be \textit{approximated} using the same approach. The key to solving such problems is to define an oracle\index{Oracle} that attributes a \textit{score}\index{Score} to a given input, rather than a yes/no answer to perfect satisfiability, and answers `yes' or `no' depending on whether that score is above some defined threshold for approximation. As an illustrative example, consider the optimisation of, say, a complex traffic network, where the goal is to maximise flow through the network. Then we might define our score to be some flow metric for the network's graph.

We then apply a Grover search repeatedly, each time incrementing this threshold until the algorithm outputs `no'. Then we know that the last input yielding the `yes' outcome had the highest score. The reason this approach is \textit{approximate} rather than \textit{exact} is that defining such a score-oracle mightn't be always efficiently implemented, or maybe it mightn't make sense at all to define score measures for a given problem. Alternately, maybe a particular problem inherently requires perfect satisfiability, and any imperfect approximation is insufficient -- if even a single constraint isn't satisfied, the system is broken.

An intuitive example of a problem taken from a far harder complexity class than \textbf{NP}-complete, that can be approximately optimised using the Grover approach is the game of chess\index{Chess}. Generalisation of the rules of chess to an \mbox{$n\times n$} board yields a game proven to reside in \textbf{EXP}-complete\index{\textbf{EXP}-complete} -- an \textit{extremely} hard class.

How do current classical computers play chess? They construct a \textit{search tree}\index{Search tree} (see Fig.~\ref{fig:search_tree}), which is explored exhaustively up to a given depth (bounded by computational resources), and assign a score to each board position, indicating the relative strength of its position. The branch in the tree with the highest scores determines the next move.

\begin{figure}[!htb]
	\includegraphics[width=0.47\textwidth]{search_tree}
	\caption{A search tree\index{Search tree} is an outwardly-directed tree graph in which each vertex represents a state of the system (i.e `board position' in chess). Edges represent actions (or `moves') available from the position from which the edge emanates. The root node represents the current position, and the depth of a vertex (i.e it's distance from the root node) represents its future time. A path through the tree therefore represents a sequence of moves up to some point in future time. If the tree had maximum depth, a route (from root to leaf) would represent an entire single game, and the complete set of routes would represent all possible ways in which the game can be played. A complete analysis of a complete search tree with maximum depth would therefore allow us to exhaustively analyse all possible ways in which a game could be played to completion, enabling optimal gameplay. However, since the number of vertices/edges in a tree graph grows exponentially with depth, the configuration search space in a game grows exponentially with how may moves into the future we wish to explore, thereby strongly limiting tree depth in practical implementations.}\label{fig:search_tree}
\end{figure}

@@

%\subsection{\textbf{BQP} optimisation problems}\index{\textbf{BQP} optimisation problems}

%The optimisation problems discussed until now have been satisfiability problems residing in \textbf{NP}-complete, limited to only quadratic quantum enhancement. Are there any other optimisation problems that natively reside in \textbf{BQP}, and therefore inherently offer exponential quantum speedup? There are, although the examples are far less ubiquitous than the plethora of satisfiability-based optimisation problems.

%\comment{To do!}

@@

While the dream-goals of perfectly optimised resource allocation and perfect economic efficiency are fantasy, we should not underestimate the impact that even modest enhancements in the efficiency of large-scale economic systems could have. While quantum computers will never perfectly solve humanity's immense resource allocation problems, even the more modest improvements they offer could be of immense value to the world economy, with enormous implications monetarily, socially, and environmentally. Indeed, optimisation problems are so ubiquitous that pursuing quantum-enhanced optimisation will be of central importance in the quantum era.
